Aggregated Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2017-10-19 aggregation/agg-JIE.tex
New foreground graph
Work on foreground + disclosure
Work on study review

\title{Aggregated Product System Models in Life Cycle Assessment: Transparency versus Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

%\usepackage{lineno}
\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

  \input{agg-intro}

  \input{agg-objective}

  \input{agg-basic-lca}

  \input{agg-lca-study}

>>>>

\title{Aggregated Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy}
\author{Brandon Kuczenski\\
Institute for Social, Behavioral, and Economic Research\\
University of California, Santa Barbara}
\date{\mydate\today}

%\usepackage{lineno}
%\let\showgrid\relax
\let\oldtimes\times
\let\times\cdot

@@

  \input{agg-intro}  % motivations

\section{Formalizing LCA Studies}

  \input{agg-objective}  % Authors, Readers, review objectives, disclosure constraints

  \input{agg-basic-lca}  % basic Leontief model

  \input{agg-lca-study}  % foreground JLCA model

  \input{agg-disclosure} % disclosure requirements satisfied

\section{Reviewing Product System Models}

Objectives in critical review vary with the study and the reviewer, but they generally include consistency with the ISO 14044 standard, the validity of methods, appropriateness of data used, correctness of interpretations, and transparency of reporting \citep[Section 6.1]{iso14044}.  Although the object of critical review is typically a written report, the use of a structured disclosure of study design introduces the possibility that reviewers can inspect the model itself and even verify the computations and reported results.  This section discusses how a disclosure in the format described above can enable robust review of a product system model.  Note that, for purposes of critical review, the objective for data privacy can be omitted because the reviewer may be granted privileged access to confidential data.

  \input{agg-foreground} % Af and graph/matrix equivalency

  \input{agg-deps}       % Ad and Bf

\section{Protecting Privacy through Aggregation}

  \input{agg-agg}        % notation for full aggregation

%  \input{agg-pagg}       % notation for partial aggregation

%  \input{agg-forms}      % the elusive Figure 4

\section{Discussion}

%  \input{agg-review}     % review objectives

%  \input{agg-privacy}    % disclosure constraints

%  \input{agg-semantic}   % the need for stable references

\section{Conclusion}

### 2017-10-19 aggregation/agg-disclosure.tex
Work on study review

This set of three submatrices can be used to represent the product system model; thus, the problem of how to disclose an LCA study design is reduced to the problem of how to communicate the contents of these submatrices accurately.  Recognizing that all three matrices are likely to be sparse (that is, that most of the entries are zero), the most efficient disclosure would be to only report non-zero entries.  A sparse matrix can be represented by a list of 3-tuples indicating (row, column, value) for each non-zero entry.  Thus, a formal representation of the PSM can be stated in six parts:
\begin{enumerate}[label=\alph*.]
\item\label{itm:fg} An ordered list of foreground nodes (rows/columns of $A_f$);
\item\label{itm:bg} An ordered list of background flows, making reference to a particular background database (rows of $A_d$, mapping to columns of $B_x$);
\item\label{itm:em} An ordered list of exchanges with environment (rows of $B_f$, mapping to entries in $\mathbf{e}$);
\item A list of 3-tuples for nonzero elements of $A_f$, in which the row and column are indices into Item~\ref{itm:fg};
\item a list of 3-tuples for nonzero elements of $A_d$, in which the rows are indices into Item~\ref{itm:bg} and the columns are indices into Item~\ref{itm:fg};
\item a list of 3-tuples for nonzero elements of $B_f$, in which the rows are indices into Item~\ref{itm:em} and the columns are indices into Item~\ref{itm:fg}.
\end{enumerate}

For a disclosure that includes the computation of an LCIA indicator or indicators, a seventh part is required:
\begin{enumerate}[resume]
\item the identity of the LCIA indicator or indicators computed (identity of $\mathbf{e}$).
\end{enumerate}

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the paper):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references to $B_x$ and $\mathbf{e}$;
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes only information for which the study is the \textbf{primary} source.
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows.  If LCIA indicator results are included, the author must also unambiguously identify the method computed.  Implications of this requirement are discussed below.

For each row in $A_d$ containing a non-zero entry, the author must unambiguously identify the version of the database used, as well as the exact process and reference flow used from the background database; the dimension (reference unit) of each reference flow must be specified; and the sign of the numeric entry in $A_d$ must be consistent with the implementation of the process in the background database.  For the reader's part, the reader must be able to obtain either the column of $B_x$, or the scalar value that

Similarly, for each row in $B_f$ containing a non-zero entry, the author must unambiguously identify the substance being exchanged with the environment, the compartment or context into which it is being exchanged, and

\subsection{Foreground Aggregation}

Often the contents of the foreground

An important intermediate result is the vector of activity levels for the foreground nodes $\tilde{\mathbf{x}}$, also visualized in Figure~\ref{fig:foreground}:
\begin{equation}
\tilde{\mathbf{x}} = (I - A_f)^{-1} \times \tilde{\mathbf{y}}
\label{eqn:inv}
\end{equation}
This vector can be used to describe an \emph{aggregated foreground}, which generates the same results as the fully expanded foreground:
\begin{equation}
\begin{array}{rl}
    \tilde{\mathbf{b}}_f & = B_f \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{a}}_d & = A_d \times \tilde{\mathbf{x}} \\
    \tilde{\mathbf{b}}_x & = B_x \times \tilde{\mathbf{a}}_d
\end{array}
\label{eqn:agg}
\end{equation}

These vectors summarize the contents of the foreground without disclosing its detailed structure.  The aggregated dependency vector $\tilde{\mathbf{a}}_d$ has the same dimension as the background database, and the aggregated emission vector $\tilde{\mathbf{b}}_f $has the same dimension as the elementary flow matrix.  The dependency vector can be transformed into an elementary flow vector by multiplying by $B_x$.  The result of the study $s$ is the sum of foreground and background impact scores:

\begin{equation}
\begin{array}{rl}
   \tilde{\mathbf{b}} & = \tilde{\mathbf{b}}_f + \tilde{\mathbf{b}}_x \\
   s &= \mathbf{e}^T \times \tilde{\mathbf{b}} \\
   & = \mathbf{e}^T \times \tilde{\mathbf{b}}_f + \mathbf{e}^T \times \tilde{\mathbf{b}}_x \\
   &= \tilde{s}_f + \tilde{s}_x
\end{array}
\label{eqn:lci}
\end{equation}

\subsubsection{Model fragments and subsystem boundaries}

The nodes that make up the foreground of an LCA study describe an implicit system boundary

Most complex

\section{Preparing and Reviewing Aggregation Results}

>>>>

This set of three submatrices can be used to represent the PSM; thus, the problem of how to disclose an LCA study design is reduced to the problem of how to communicate the contents of these submatrices accurately.  Recognizing that all three matrices are likely to be sparse (that is, that most of the entries are zero), the most efficient disclosure would be to only report non-zero entries.  A sparse matrix can be represented by a list of 3-tuples indicating (row, column, value) for each non-zero entry.  Thus, a formal disclosure of the PSM can be stated in six parts:
\begin{enumerate}[label={\em d-\roman*}., ref={\em d-\roman*}]
\item\label{itm:fg} An ordered list of foreground nodes (rows/columns of $A_f$);
\item\label{itm:bg} An ordered list of background flows, making reference to a particular background database (rows of $A_d$, mapping to columns of $B_x$);
\item\label{itm:em} An ordered list of exchanges with environment (rows of $B_f$, mapping to entries in $\mathbf{e}$);
\item\label{itm:af} A list of 3-tuples for nonzero elements of $A_f$, in which the row and column are indices into Item~\ref{itm:fg};
\item\label{itm:ad} a list of 3-tuples for nonzero elements of $A_d$, in which the rows are indices into Item~\ref{itm:bg} and the columns are indices into Item~\ref{itm:fg};
\item\label{itm:bf} a list of 3-tuples for nonzero elements of $B_f$, in which the rows are indices into Item~\ref{itm:em} and the columns are indices into Item~\ref{itm:fg}.
\end{enumerate}

A disclosure containing these six elements satisfies most of the requirements established at the top of this section (the exception is ensuring the privacy of confidential data, which is treated later in the article):
\begin{itemize}
  \item The result is \textbf{computable} using Eq.~\ref{eqn:study};
  \item the computation is \textbf{complete} and \textbf{minimal} because all components are required and no superfluous components are included;
  \item The results are \textbf{reproducible} if the reader is able to accurately interpret the author's references to $B_x$ and $\mathbf{e}$;
  \item The disclosure includes only information for which the author can claim \textbf{authority};
  \item The disclosure includes only information for which the study is the \textbf{primary} source.
\end{itemize}
As mentioned, the proper interpretation of the disclosure requires that the reader accurately interprets the author's references to background databases and elementary flows, including mutually consistent implementation of the LCIA vector $\mathbf{e}$.  These aspects are out of scope for the purposes of this paper, and we will assume that a shared understanding between author and reader can be reached.  However, the the challenges of operationalizing this requirement are tremendously complex.  Some implications of the requirement for shared agreement are considered in the Discussion.

### 2017-10-19 aggregation/agg-foreground.tex
Work on foreground + disclosure
Work on study review

\subsection{The Foreground}

The foreground of an LCA study is a collection of nodes that represent points in the inventory model where flows are exchanged.  The foreground is modelled as a weighted directed graph in which there is a 1:1 correspondence between nodes and product flows.

The foreground matrix $A_f$ reports the adjacency and weights of edges in the graph, and thus describes the relationships among these nodes.  These nodes typically represent unit processes, but they can also indicate exchanges across an implicit system boundary that encloses the nodes.  If a column in $A_f$ is empty, then that is an indication that the corresponding product flow is excluded from the foreground.  In order for this flow to be accounted for in the model, it must either be ``terminated'' in the dependency matrix $A_d$, or represented in the corresponding column of $B_f$ as an emission.  Flows whose entire column consists of 0s throughout $A_f$, $A_d$, and $B_f$ are ``cut offs'' that exit the model with zero burdens.

@@

Generally, the foreground matrix $A_f$ can represent any process-flow model with a 1:1 correspondence between processes and flows, i.e. any allocated supply-and-use table, but some fundamental designs are common.  Some simple foreground models are shown in Figure~\ref{fig:fragments}.  The first (a) is a sequential model, in which each node requires one foreground input and generates one output. This model is equivalent to a ``gate to gate'' model.  Here the weights $k_i$ indicate the amount of the preceding reference flow that is required by the subsequent node.  Figure~\ref{fig:fragments}(b) shows an additive model, in which the outputs of several foreground nodes are added together, equivalent to a ``mixer'' or a horizontal average.  In this arrangement the weights %represent the relative weights of each input, and
should add up to a unit output of the reference node.

add a cyclic model

\subsubsection{Model fragments and subsystem boundaries}

The nodes that make up the foreground of an LCA study describe an implicit system boundary

Most complex

\section{Preparing and Reviewing Aggregation Results}

>>>>

\subsection{Interpreting the Foreground}

The foreground of an LCA study is a collection of nodes that represent points in the inventory model where flows are exchanged.  The foreground matrix $A_f$ describes in precise terms what components are included in the product system and how they connect to one another.  The foreground is modelled as a weighted directed graph in which there is a 1:1 correspondence between nodes and product flows.  The foreground matrix $A_f$ reports the adjacency and weights of edges in the graph, and thus describes the relationships among these nodes.  These nodes typically represent unit processes, but they can also indicate exchanges across an implicit system boundary that encloses the foreground.

A reviewer with access to items \ref{itm:fg} and \ref{itm:af} in the disclosure would be able to construct $A_f$ and automatically create a process-flow diagram, annotated with information about each node.  Generally, the foreground matrix $A_f$ can represent any process-flow model with a 1:1 correspondence between processes and flows, i.e. any allocated supply-and-use table, but some fundamental designs are common.  Some simple foreground models are shown in Figure~\ref{fig:fragments}.

@@

The first (a) is a sequential model, in which each node requires one foreground input and generates one output. This model is equivalent to a ``gate to gate'' model.  Here the weights $k_i$ indicate the amount of the preceding reference flow that is required by the subsequent node.  Figure~\ref{fig:fragments}(b) shows an additive model, in which the outputs of several foreground nodes are added together, equivalent to a ``mixer'' or a horizontal average.  In this arrangement the weights %represent the relative weights of each input, and
should add up to a unit output of the reference node.  Finally, Figure~\ref{fig:fragments}(c) shows a foreground model with a cyclic dependency, where some of the reference output is consumed by another foreground node.

A typical PSM may contain multiple fragmentary elements that are interconnected.  An example of foreground composed of several fragments is illustrated in Figure~\ref{fig:fragments}(d).  Here, the nodes labeled 1--5 represent one fragment, which generates the foreground's canonical reference flow $\tilde{y}$. This fragment requires two interior flows from separate fragments ($y_0$ and $y_1$), and has two unconnected flows (4 and 5).  The reference $y_0$ is supplied by a second fragment, made up of nodes 6-8.  The reference $y_1$ is supplied by another fragment made of only one node (9). The reference flow $y_1$ is consumed in two different places by the other fragments.
